Edge sign prediction in general graphs and Correlation Clustering

### 2017-10-24 edge/experiments.tex
take Marc feedback on the edge model into account
- split into more subsection
- add an example
- describe and name the methods earlier
- add a paragraph comparing all the methods
- make it more clear there will be only synthetic data

\section{Experimental analysis}
\label{sec:edge_exp}

We start by extracting from the conceptual discussion of the previous section the five methods we
compare in our experiments. They all take as input an attributed graph, \ie{} $G=(V, E)$ and $X$.
For evaluation purposes, we describe how to extract from their output a set of $k$ directions and an
assignment $\yhat$ of every edge to one of these directions (that is $\yhat$ is a function from $E$
to $\rangesk$ such that $\yhat_{uv}$ is the index of the direction assigned to the edge $(u,v)$).

\begin{enumerate}[1.]

  \item Our baseline is a \kmeans{} clustering of the normalized $s_{uv}$ edge similarity
    vectors. This naturally provides a \yhat{} assignment, and the directions are the normalized
    cluster centers.

  \item Our first method is a \lloyd{}-like iterative refinement of the $k$-means solution, where we
    alternatively reassign edges based on their score and recompute the directions to maximize those
    scores.

  \item Then we consider the \textsc{Combined} optimization of the two edge and node terms of
    \eqref{eq:edge_full}, which directly provides a set of directions. The assignment is then simply
    the direction that gives the highest goodness.

  \item As for matrix methods, we first do a \textsc{Frank--Wolfe} maximization of
    \eqref{eq:edge_FW} over the $\delta$-ball of the nuclear norm. Afterwards, we perform a
    \kmeans{} clustering of its columns and use it as assignment, while the directions are again the
    normalized cluster centers.

  \item \textsc{Explicit} low-rank $PQ^T$ factorization, initialized with the solution of
    \combined{} (that is we set $P_0 \in \mathbb{M}^{d\times k}$ to be the final directions found by
    \combined{} and the rows of $Q_0 \in \mathbb{M}^{|E|\times k}$ to be equal to the corresponding
    $e_{y_{uv}}$) and minimizing the same objective of \eqref{eq:edge_PQ}. We take the directions to
    be the columns of the resulting $P$, and the assignment to be a \kmeans{} clustering of the rows
    of $Q$.
\end{enumerate}

\subsection{Synthetic data}
\label{ssec:edge_synth}

We try those five methods on data generated synthetically to fit our model describe previously and
study how they perform under various conditions.

\paragraph{Generation}
\label{par:edge_generation}

@@

We first generate the $k$ unit-norm directions $\mathcal{D}_k=\{w_1, \ldots, w_k\}$, independently
of the graph. Furthermore, we control the number of coordinates $n_o$ that overlap with each other.
That is, when $n_o = 0$, each direction has exactly \nicefrac{k}{d}\footnote{We choose $d$ to be a
multiple of $k$.} non-zero dimension that are disjoint. On the other hand, if for instance $n_o=5$,
then $5$ of the $d$ dimensions will have two directions with non-zero components on it. Those
non-zero components are drawn \uar{} from $[-1, 1]$ and the each directions is then normalized. For
instance, with $d=6$, $k=3$ and $n_o=2$, we could generate the following directions, where the dots
represents non-zero numbers and the red ones are overlapping:
\vspace{-.5\baselineskip}
\begin{center}
  \begin{tabular}{ccc}
    $w_1$ & $w_2$ & $w_3$ \\
    $\begin{pmatrix}
      \bullet \\
      0 \\
      \bullet \\
      0 \\
      0 \\
      \textcolor{Red}{\bullet}
    \end{pmatrix}$ &
    $\begin{pmatrix}
      0 \\
      \bullet \\
      0 \\
      \textcolor{Red}{\bullet} \\
      \bullet \\
      0
    \end{pmatrix}$ &
    $\begin{pmatrix}
      0 \\
      0 \\
      0 \\
      \bullet \\
      0 \\
      \bullet
    \end{pmatrix}$
  \end{tabular}
\end{center}

@@

\paragraph{Results}
\label{par:edge_synth_results}

>>>>

\section{Synthetic experiments}
\label{sec:edge_exp}

In the previous section, we introduced five methods to address \autoref{p:edge_full}, which we now
compare in our experiments. Specifically, this includes:
\begin{enumerate}[1.,nosep]
\item the \kmeans{} baseline,
\item our \lloyd{}-like iterative refinement,
\item the \combined{} convex optimization of \eqref{eq:edge_soft},
\item the \fwa{} convex optimization of \eqref{eq:edge_FW} and
\item the \pqt{} explicit low-rank factorization, optimizing \eqref{eq:edge_PQ}.
\end{enumerate}
They all take as input a node-attributed graph, that is a pair $G=(V, E)$ and $X$.
For evaluation purposes, we extract from their output a set of $k$ directions and an
assignment of every edge to one of these directions.
We try those five methods on data generated synthetically to fit our model described previously in
\autoref{sub:edge_setting}, and study how they perform under various conditions.

% (that is $\yhat$ is a function from $E$
% to $\rangesk$ such that $\yhat_{uv}$ is the index of the direction assigned to the edge $(u,v)$).

\subsection{Generation}
\label{sub:edge_generation}

@@

We first generate the $k$ unit-norm directions $\mathcal{D}_k=\{w_1, \ldots, w_k\}$, independently
of the graph. Furthermore, we control the number of coordinates $n_o$ that overlap with each other.
That is, when $n_o = 0$, each direction has exactly \nicefrac{k}{d}\footnote{We choose $d$ to be a
multiple of $k$.} non-zero dimension that are disjoint. On the other hand, if for instance $n_o=5$,
then $5$ of the $d$ dimensions will have two directions with non-zero components on it. Those
non-zero components are drawn \uar{} from $[-1, 1]$ and the each directions is then normalized. For
instance, with $d=6$, $k=3$ and $n_o=2$, we could generate the following directions, where the dots
represents non-zero numbers and the blue ones are overlapping:
\vspace{-.5\baselineskip}
\begin{center}
  \begin{tabular}{ccc}
    $w_1$ & $w_2$ & $w_3$ \\
    $\begin{pmatrix}
      \bullet \\
      0 \\
      \bullet \\
      0 \\
      0 \\
      \textcolor{DodgerBlue}{\bullet}
    \end{pmatrix}$ &
    $\begin{pmatrix}
      0 \\
      \bullet \\
      0 \\
      \textcolor{DodgerBlue}{\bullet} \\
      \bullet \\
      0
    \end{pmatrix}$ &
    $\begin{pmatrix}
      0 \\
      0 \\
      0 \\
      \bullet \\
      0 \\
      \bullet
    \end{pmatrix}$
  \end{tabular}
\end{center}

@@

\subsection{Results}
\label{sub:edge_synth_results}
